The transboundary nature of the world’s exploited marine species

### 2021-03-19 Scripts/Manuscript_Draft.Rmd
added high seas descrete stocks

Distributions of marine species around the world are not constrained by human-made boundaries; rather they are shaped by biotic and abiotic factors as well as species' evolutionary history [@Hutchinson:1957nt; @Nelson:2016kk]. A species can be widely distributed (cosmopolitan) or endemic [@Nelson:2016kk]. Fisheries management is predicated on the definition of “stocks”, delineated, in most cases by human-made spatial boundaries that often do not correspond to biologically-meaningful population units [@Song:2017iua; @FredstonHermann:2018kp]. The development of Exclusive Economic Zones (EEZs) under Part V of the United Nations Convention on the Law of the Seas (UNCLOS) in the early 80s [@UnitedNations:1986tl], for instance, extended political boundaries from 12 to 200 nautical miles to give coastal nations property rights over marine resources [@osthagen:2020hd]. However, EEZ boundaries cut across the distribution of many species, creating shared stocks between nations [@UnitedNations:1986tl]. Shared stocks can be classified into three non-exclusive categories; "transboundary" - stocks that cross the EEZs of two or more bordering coastal states; "straddling" - stocks that cross neighboring EEZs and the adjacent high seas; and "highly migratory" - stocks that cross non-neighboring EEZs and the high seas (mainly tunas) [@Munro:2004th]. My study focuses on the "transboundary" nature of shared species exploited by fisheries operating within EEZs.

>>>>

Distributions of marine species around the world are not constrained by human-made boundaries; rather they are shaped by biotic and abiotic factors as well as species' evolutionary history [@Hutchinson:1957nt; @Nelson:2016kk]. A species can be widely distributed (cosmopolitan) or endemic [@Nelson:2016kk]. Fisheries management is predicated on the definition of “stocks”, delineated, in most cases by human-made spatial boundaries that often do not correspond to biologically-meaningful population units [@Song:2017iua; @FredstonHermann:2018kp]. The development of Exclusive Economic Zones (EEZs) under Part V of the United Nations Convention on the Law of the Seas (UNCLOS) in the early 80s [@UnitedNations:1986tl], for instance, extended political boundaries from 12 to 200 nautical miles to give coastal nations property rights over marine resources [@osthagen:2020hd]. However, EEZ boundaries cut across the distribution of many species, creating shared stocks between nations [@UnitedNations:1986tl]. Shared stocks can be classified into four non-exclusive categories; "transboundary" - stocks that cross the EEZs of two or more bordering coastal states; "straddling" - stocks that cross neighboring EEZs and the adjacent high seas; "highly migratory" - stocks that cross non-neighboring EEZs and the high seas (mainly tunas); and finally "discrete" - stocks that are only present in the high seas [@Munro:2004th]. My study focuses on the "transboundary" nature of shared species exploited by fisheries operating within EEZs.
